Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach

### 2013-10-04 inst/doc/nonparametric-bayes.tex
compile draft

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{color}
\usepackage{fancyvrb}
\DefineShortVerb[commandchars=\\\{\}]{\|}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

\section{Abstract}

@@

\section{Introduction}

@@

The sudden collapse of fisheries and other ecosystems is an increasingly
widespread phenomenon and a pressing concern for ecological management
and conservation. Ecological dynamics are frequently complex and
difficult to measure, making uncertainty in our understanding a
prediction a persistent challenge to effective management.
Decision-theoretic approaches provide a framework to determine the best
sequence of actions in face of uncertainty, but only when that
uncertainty can be meaningfully quantified {[}@Fischer2009{]}.

Uncertainty enters the decision-making process at many levels: intrinsic
stochasticity in biological processes, measurements, and implementation
of policy
\citetext{\citealp[\emph{e.g.}][]{Reed1979}; \citealp{Clark1986}; \citealp[@Roughgarden1997;][]{Sethi2005}},
parameteric uncertainty
\citetext{\citealp[\emph{e.g.}][]{Ludwig1982}; \citealp[@Hilborn1997;
@McAllister1998;][]{Schapaugh2013}}, and model or structural uncertainty
\citep[\emph{e.g.}][]{Williams2001, Cressie2009, Athanassoglou2012}. Of
these, structural uncertainty incorporates the least a priori knowledge
or assumptions and is generally the hardest to quantify. Typical
approaches assume a weak notion of uncertainty where a correct or
reasonable approximation of the dynamics must be identified from among a
handful of alternative models. Here we consider an approach that
addresses uncertainty at each of these levels without assuming the
dynamics follow a particular (i.e.~parametric) structure.

An additional source of uncertainty that has recieved less
attention\footnote{The concept of adaptive probing is one area that has
  explicitly addressed this kind of uncertainty, reaching rather
  opposite conclusions than what we observe here. Adaptive probing
  strategies follow from ``Dual Control'' or ``Active Adaptive
  Management'' approaches (e.g. \citet{Ludwig1982}) that can trade off
  short term utility by choosing actions that can reduce uncertainty.
  Adaptive probing strategies arise when it is valuable to intentionally
  force a system far from the observed values even when the expected
  value such actions is low, as it provides much faster learning and
  consequent reduction of model uncertainty that can allow greater value
  to be derived later on. For instance, \citet{Ludwig1982} show that it
  may be advantageous to fish an unexploited population very heavily at
  first to obtain a better estimate of the recruitment rate. This
  intuitive strategy when a population is governed by a Ricker or
  Beverton-Holt-like dynamic would clearly be disastrous if instead the
  dynamics contained an unforeseen tipping point. The best way to learn
  where the edge lies may be to walk up to it, but it is also the most
  dangerous.} arises when applying a dynamical model outside the range
of data on which it has been estimated. This extrapolation uncertainty
is felt most keenly in decision-theoretic applications, as (a) exploring
the potential action space typically involves considering actions that
may move the system outside the range of observed behavior, and (b)
decision-theoretic alogrithms rely not only on reasonable estimates of
the expected outcomes, but depend on the weights given to all possible
outcomes \citep[\emph{e.g.}][]{Weitzman2013}.

Concerns over the potential for tipping points in ecological dynamics
\citep{Scheffer2001} highlight the dangers of uncertainty in ecological
management and pose a substantial challenge to existing
decision-theoretic approaches \citep{Brozovic2011}. Because intervention
is often too late after a tipping point has been crossed (but see
\citet{Hughes2013}), management is most often concerned with avoiding
potentially catastrophic tipping points before any data is available at
or following a transition that would more clearly reveal these regime
shift dynamics \citep[e.g.][]{Bestelmeyer2012}.

@@

Here we illustrate how a stochastic dynamic programming (SDP) algorithm
{[}\citet{Mangel1988}; @Possingham1997, Marescot2013{]} can be driven by
the predictions from a Bayesian non-parametric (BNP) approach
\citep{Munch2005a}.\\This provides two distinct advantages compared with
contemporary approaches. First, using a BNP sidesteps the need for an
accurate model-based description of the system dynamics.\\Second, the
BNP can better reflect uncertainty that arises when extrapolating a
model outside of the data on which it was fit.\\We illustrate that when
the correct model is not known, this latter feature is crucial to
providing a robust decision-theoretic approach in face of substantial
structural uncertainty.

\subsection{Two central challenges:}

@@

\subsection{Why we don't know the model}

\begin{itemize}
\item
  Complex dynamics \citep{Glaser2013}(http://doi.org/10.1111/faf.12037
  ``Complex dynamics may limit prediction in marine fisheries'')
\item
  model choice and model averaging approaches
\end{itemize}

\subsection{Why we don't have data where we need it}

@@

The economic value and ecological concern have made marine fisheries the
crucible for much of the founding work
\citep{Gordon1954, Reed1979, May1979, Ludwig1982} in managing ecosystems
under uncertainty. Global trends \citep{Worm2006} and controversy
\citep{Hilborn2007, Worm2009} have made understanding these challenges
all the more pressing.

@@

The nature of decision-making problems provides a convenient way to
compare models. Rather than compare models in terms of best fit to data
or fret over the appropriate penalty for model complexity, model
performance is defined in the concrete terms of the decision-maker's
objective function, which we will take as given. (Much argument can be
made over the `correct' objective function, e.g.~how to account for the
social value of fish left in the sea vs.~the commercial value of fish
harvested; see \citet{Halpern2013} for further discussion of this issue.
Alternatively, we can always compare model performance across multiple
potential objective functions.) The decision-maker does not necessarily
need a model that provides the best mechanistic understanding or the
best long-term outcome, but rather the one that best estimates the
probabilities of being in different states as a result of the possible
actions.

\subsection{Background on the Gaussian Process}

@@

The use of Gaussian process (GP) regression (or ``kriging'' in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems \citep{Kocijan2005}, and
was first introduced in the context ecological modeling and fisheries
management in \citet{Munch2005}. An accessible and thorough introduction
to the formulation and use of GPs can be found in \citet{Rasmussen2006}.

@@

The posterior distribution for the hyper-parameters of the Gaussian
process model are estimated by Metropolis-Hastings algorithm, again with
details and code provided in the Appendix. \citet{Rasmussen2006}
provides an excellent general introduction to Gaussian Processes and
\citet{Munch2005} first discusses their application in the context of
population dynamics models such as fisheries stock-recruitment
relationships.

\section{Approach and Methods}

\subsubsection{Summary of approach}

\subsubsection{Statement of the optimal control problem}

@@

\subsubsection{Parametric models}

@@

We infer posterior distributions for the parameters of each model in a
Bayesian context using Gibbs sampling (implemented in R {[}@RTeam{]}
using jags, {[}@R2jags{]}). We choose uninformative uniform priors for
all parameters (See Appendix, Figures S1-S3, and Table S1, and the R
code provided). One-step-ahead predictions of these model fits are shown
in Figure 1.

@@

\subsubsection{The Gaussian Process model}

@@

\section{Results}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{require}\NormalTok{(MASS)}
\NormalTok{step_ahead <- function(x, f, p)\{}
  \NormalTok{h = }\DecValTok{0}
  \NormalTok{x_predict <- }\KeywordTok{sapply}\NormalTok{(x, f, h, p)}
  \NormalTok{n <- }\KeywordTok{length}\NormalTok{(x_predict) - }\DecValTok{1}
  \NormalTok{y <- }\KeywordTok{c}\NormalTok{(x[}\DecValTok{1}\NormalTok{], x_predict[}\DecValTok{1}\NormalTok{:n])}
  \NormalTok{y}
\NormalTok{\}}
\NormalTok{step_ahead_posteriors <- function(x)\{}
\NormalTok{gp_f_at_obs <- }\KeywordTok{gp_predict}\NormalTok{(gp, x, }\DataTypeTok{burnin=}\FloatTok{1e4}\NormalTok{, }\DataTypeTok{thin=}\DecValTok{300}\NormalTok{)}
\NormalTok{df_post <- }\KeywordTok{melt}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(}\KeywordTok{sample}\NormalTok{(}\DecValTok{100}\NormalTok{), }
  \NormalTok{function(i)\{}
    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{time =} \DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(x), }\DataTypeTok{stock =} \NormalTok{x, }
                \DataTypeTok{GP =} \KeywordTok{mvrnorm}\NormalTok{(}\DecValTok{1}\NormalTok{, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),}
                \DataTypeTok{True =} \KeywordTok{step_ahead}\NormalTok{(x,f,p),  }
                \DataTypeTok{MLE =} \KeywordTok{step_ahead}\NormalTok{(x,f,est$p), }
                \DataTypeTok{Allen =} \KeywordTok{step_ahead}\NormalTok{(x, allen_f, pardist[i,]), }
                \DataTypeTok{Ricker =} \KeywordTok{step_ahead}\NormalTok{(x, ricker_f, ricker_pardist[i,]), }
                \DataTypeTok{Myers =} \KeywordTok{step_ahead}\NormalTok{(x, myers_f, myers_pardist[i,]))}
  \NormalTok{\}), }\DataTypeTok{id=}\KeywordTok{c}\NormalTok{(}\StringTok{"time"}\NormalTok{, }\StringTok{"stock"}\NormalTok{))}
\NormalTok{\}}

\NormalTok{df_post <- }\KeywordTok{step_ahead_posteriors}\NormalTok{(x)}

\KeywordTok{ggplot}\NormalTok{(df_post) + }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, stock)) + }
  \KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, value, }\DataTypeTok{col=}\NormalTok{variable, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(L1,variable)), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) + }
  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{))) }
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{policies <- }\KeywordTok{melt}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{stock=}\NormalTok{x_grid, }\KeywordTok{sapply}\NormalTok{(OPT, function(x) x_grid[x])), }\DataTypeTok{id=}\StringTok{"stock"}\NormalTok{)}
\KeywordTok{names}\NormalTok{(policies) <- }\KeywordTok{c}\NormalTok{(}\StringTok{"stock"}\NormalTok{, }\StringTok{"method"}\NormalTok{, }\StringTok{"value"}\NormalTok{)}

\KeywordTok{ggplot}\NormalTok{(policies, }\KeywordTok{aes}\NormalTok{(stock, stock - value, }\DataTypeTok{color=}\NormalTok{method)) +}
  \KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{lwd=}\FloatTok{1.2}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) + }\KeywordTok{xlab}\NormalTok{(}\StringTok{"stock size"}\NormalTok{) + }\KeywordTok{ylab}\NormalTok{(}\StringTok{"escapement"}\NormalTok{)  +}
  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-Figure2.pdf}
\caption{The steady-state optimal policy (infinite boundary) calculated
under each model. Policies are shown in terms of target escapement,
$S_t$, as under models such as this a constant escapement policy is
expected to be optimal \citep{Reed1979}.}
\end{figure}

The resulting optimal management strategy based on each of the inferred
models is shown in Figure 2, against the optimal strategy given the true
underlying dynamics. Policies are shown in terms of target escapement,
$S_t$. Under models such as this a constant escapement policy is
expected to be optimal \citep{Reed1979}, whereby population levels below
a certain size $S$ are unharvested, while above that size the harvest
strategy aims to return the population to $S$, resulting in the
hockey-stick shaped policies shown.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{(sims_data) + }
  \KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, fishstock, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(reps,method), }\DataTypeTok{color=}\NormalTok{method), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) +}
  \KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Avoiding tipping points in the management of ecological systems: a non-parametric Bayesian approach},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

\section{Abstract}\label{abstract}

@@

\section{Introduction}\label{introduction}

@@

The sudden collapse of fisheries and other ecosystems is an increasingly
widespread phenomenon and a pressing concern for ecological management
and conservation. Ecological dynamics are frequently complex and
difficult to measure, making uncertainty in our understanding a
prediction a persistent challenge to effective management.
Decision-theoretic approaches provide a framework to determine the best
sequence of actions in face of uncertainty, but only when that
uncertainty can be meaningfully quantified ({\textbf{???}}).

Uncertainty enters the decision-making process at many levels: intrinsic
stochasticity in biological processes, measurements, and implementation
of policy (\emph{e.g.} Reed 1979; C. W. Clark and Kirkwood 1986;
{\textbf{???}}; Sethi et al. 2005), parameteric uncertainty (\emph{e.g.}
Ludwig and Walters 1982; {\textbf{???}}; {\textbf{???}}; Schapaugh and
Tyre 2013), and model or structural uncertainty (\emph{e.g.} B. K.
Williams 2001; Cressie et al. 2009; Athanassoglou and Xepapadeas 2012).
Of these, structural uncertainty incorporates the least a priori
knowledge or assumptions and is generally the hardest to quantify.
Typical approaches assume a weak notion of uncertainty where a correct
or reasonable approximation of the dynamics must be identified from
among a handful of alternative models. Here we consider an approach that
addresses uncertainty at each of these levels without assuming the
dynamics follow a particular (i.e.~parametric) structure.

An additional source of uncertainty that has recieved less
attention\footnote{The concept of adaptive probing is one area that has
  explicitly addressed this kind of uncertainty, reaching rather
  opposite conclusions than what we observe here. Adaptive probing
  strategies follow from ``Dual Control'' or ``Active Adaptive
  Management'' approaches (e.g. Ludwig and Walters (1982)) that can
  trade off short term utility by choosing actions that can reduce
  uncertainty. Adaptive probing strategies arise when it is valuable to
  intentionally force a system far from the observed values even when
  the expected value such actions is low, as it provides much faster
  learning and consequent reduction of model uncertainty that can allow
  greater value to be derived later on. For instance, Ludwig and Walters
  (1982) show that it may be advantageous to fish an unexploited
  population very heavily at first to obtain a better estimate of the
  recruitment rate. This intuitive strategy when a population is
  governed by a Ricker or Beverton-Holt-like dynamic would clearly be
  disastrous if instead the dynamics contained an unforeseen tipping
  point. The best way to learn where the edge lies may be to walk up to
  it, but it is also the most dangerous.} arises when applying a
dynamical model outside the range of data on which it has been
estimated. This extrapolation uncertainty is felt most keenly in
decision-theoretic applications, as (a) exploring the potential action
space typically involves considering actions that may move the system
outside the range of observed behavior, and (b) decision-theoretic
alogrithms rely not only on reasonable estimates of the expected
outcomes, but depend on the weights given to all possible outcomes
(\emph{e.g.} Weitzman 2013).

Concerns over the potential for tipping points in ecological dynamics
(Scheffer et al. 2001) highlight the dangers of uncertainty in
ecological management and pose a substantial challenge to existing
decision-theoretic approaches (Brozović and Schlenker 2011). Because
intervention is often too late after a tipping point has been crossed
(but see Hughes et al. (2013)), management is most often concerned with
avoiding potentially catastrophic tipping points before any data is
available at or following a transition that would more clearly reveal
these regime shift dynamics (e.g. Bestelmeyer et al. 2012).

@@

Here we illustrate how a stochastic dynamic programming (SDP) algorithm
(Mangel and Clark 1988; {\textbf{???}}) can be driven by the predictions
from a Bayesian non-parametric (BNP) approach (Munch, Kottas, and Mangel
2005).\\This provides two distinct advantages compared with contemporary
approaches. First, using a BNP sidesteps the need for an accurate
model-based description of the system dynamics.\\Second, the BNP can
better reflect uncertainty that arises when extrapolating a model
outside of the data on which it was fit.\\We illustrate that when the
correct model is not known, this latter feature is crucial to providing
a robust decision-theoretic approach in face of substantial structural
uncertainty.

\subsection{Two central challenges:}\label{two-central-challenges}

@@

\subsection{Why we don't know the
model}\label{why-we-dont-know-the-model}

\begin{itemize}
\item
  Complex dynamics (Glaser et al. 2013)(http://doi.org/10.1111/faf.12037
  ``Complex dynamics may limit prediction in marine fisheries'')
\item
  model choice and model averaging approaches
\end{itemize}

\subsection{Why we don't have data where we need
it}\label{why-we-dont-have-data-where-we-need-it}

@@

The economic value and ecological concern have made marine fisheries the
crucible for much of the founding work (Gordon 1954; Reed 1979; May et
al. 1979; Ludwig and Walters 1982) in managing ecosystems under
uncertainty. Global trends (Worm et al. 2006) and controversy (Hilborn
2007; Worm et al. 2009) have made understanding these challenges all the
more pressing.

@@

The nature of decision-making problems provides a convenient way to
compare models. Rather than compare models in terms of best fit to data
or fret over the appropriate penalty for model complexity, model
performance is defined in the concrete terms of the decision-maker's
objective function, which we will take as given. (Much argument can be
made over the `correct' objective function, e.g.~how to account for the
social value of fish left in the sea vs.~the commercial value of fish
harvested; see Halpern et al. (2013) for further discussion of this
issue. Alternatively, we can always compare model performance across
multiple potential objective functions.) The decision-maker does not
necessarily need a model that provides the best mechanistic
understanding or the best long-term outcome, but rather the one that
best estimates the probabilities of being in different states as a
result of the possible actions.

\subsection{Background on the Gaussian
Process}\label{background-on-the-gaussian-process}

@@

The use of Gaussian process (GP) regression (or ``kriging'' in the
geospatial literature) to formulate a predictive model is relatively new
in the context of modeling dynamical systems (Kocijan et al. 2005), and
was first introduced in the context ecological modeling and fisheries
management in Munch et al. (2005). An accessible and thorough
introduction to the formulation and use of GPs can be found in Rasmussen
and Williams (2006).

@@

The posterior distribution for the hyper-parameters of the Gaussian
process model are estimated by Metropolis-Hastings algorithm, again with
details and code provided in the Appendix. Rasmussen and Williams (2006)
provides an excellent general introduction to Gaussian Processes and
Munch et al. (2005) first discusses their application in the context of
population dynamics models such as fisheries stock-recruitment
relationships.

\section{Approach and Methods}\label{approach-and-methods}

\subsubsection{Summary of approach}\label{summary-of-approach}

\subsubsection{Statement of the optimal control
problem}\label{statement-of-the-optimal-control-problem}

@@

\subsubsection{Parametric models}\label{parametric-models}

@@

We infer posterior distributions for the parameters of each model in a
Bayesian context using Gibbs sampling (implemented in R ({\textbf{???}})
using jags, ({\textbf{???}})). We choose uninformative uniform priors
for all parameters (See Appendix, Figures S1-S3, and Table S1, and the R
code provided). One-step-ahead predictions of these model fits are shown
in Figure 1.

@@

\subsubsection{The Gaussian Process
model}\label{the-gaussian-process-model}

@@

\section{Results}\label{results}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{require}\NormalTok{(MASS)}
\NormalTok{step_ahead <-}\StringTok{ }\NormalTok{function(x, f, p)\{}
  \NormalTok{h =}\StringTok{ }\DecValTok{0}
  \NormalTok{x_predict <-}\StringTok{ }\KeywordTok{sapply}\NormalTok{(x, f, h, p)}
  \NormalTok{n <-}\StringTok{ }\KeywordTok{length}\NormalTok{(x_predict) -}\StringTok{ }\DecValTok{1}
  \NormalTok{y <-}\StringTok{ }\KeywordTok{c}\NormalTok{(x[}\DecValTok{1}\NormalTok{], x_predict[}\DecValTok{1}\NormalTok{:n])}
  \NormalTok{y}
\NormalTok{\}}
\NormalTok{step_ahead_posteriors <-}\StringTok{ }\NormalTok{function(x)\{}
\NormalTok{gp_f_at_obs <-}\StringTok{ }\KeywordTok{gp_predict}\NormalTok{(gp, x, }\DataTypeTok{burnin=}\FloatTok{1e4}\NormalTok{, }\DataTypeTok{thin=}\DecValTok{300}\NormalTok{)}
\NormalTok{df_post <-}\StringTok{ }\KeywordTok{melt}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(}\KeywordTok{sample}\NormalTok{(}\DecValTok{100}\NormalTok{), }
  \NormalTok{function(i)\{}
    \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{time =} \DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(x), }\DataTypeTok{stock =} \NormalTok{x, }
                \DataTypeTok{GP =} \KeywordTok{mvrnorm}\NormalTok{(}\DecValTok{1}\NormalTok{, gp_f_at_obs$Ef_posterior[,i], gp_f_at_obs$Cf_posterior[[i]]),}
                \DataTypeTok{True =} \KeywordTok{step_ahead}\NormalTok{(x,f,p),  }
                \DataTypeTok{MLE =} \KeywordTok{step_ahead}\NormalTok{(x,f,est$p), }
                \DataTypeTok{Allen =} \KeywordTok{step_ahead}\NormalTok{(x, allen_f, pardist[i,]), }
                \DataTypeTok{Ricker =} \KeywordTok{step_ahead}\NormalTok{(x, ricker_f, ricker_pardist[i,]), }
                \DataTypeTok{Myers =} \KeywordTok{step_ahead}\NormalTok{(x, myers_f, myers_pardist[i,]))}
  \NormalTok{\}), }\DataTypeTok{id=}\KeywordTok{c}\NormalTok{(}\StringTok{"time"}\NormalTok{, }\StringTok{"stock"}\NormalTok{))}
\NormalTok{\}}

\NormalTok{df_post <-}\StringTok{ }\KeywordTok{step_ahead_posteriors}\NormalTok{(x)}

\KeywordTok{ggplot}\NormalTok{(df_post) +}\StringTok{ }\KeywordTok{geom_point}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, stock)) +}\StringTok{ }
\StringTok{  }\KeywordTok{geom_line}\NormalTok{(}\KeywordTok{aes}\NormalTok{(time, value, }\DataTypeTok{col=}\NormalTok{variable, }\DataTypeTok{group=}\KeywordTok{interaction}\NormalTok{(L1,variable)), }\DataTypeTok{alpha=}\NormalTok{.}\DecValTok{1}\NormalTok{) +}\StringTok{ }
\StringTok{  }\KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey, }\DataTypeTok{guide =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{override.aes =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{1}\NormalTok{))) }
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{policies <-}\StringTok{ }\KeywordTok{melt}\NormalTok{(}\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{stock=}\NormalTok{x_grid, }\KeywordTok{sapply}\NormalTok{(OPT, function(x) x_grid[x])), }\DataTypeTok{id=}\StringTok{"stock"}\NormalTok{)}
\KeywordTok{names}\NormalTok{(policies) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"stock"}\NormalTok{, }\StringTok{"method"}\NormalTok{, }\StringTok{"value"}\NormalTok{)}

\KeywordTok{ggplot}\NormalTok{(policies, }\KeywordTok{aes}\NormalTok{(stock, stock -}\StringTok{ }\NormalTok{value, }\DataTypeTok{color=}\NormalTok{method)) +}
\StringTok{  }\KeywordTok{geom_line}\NormalTok{(}\DataTypeTok{lwd=}\FloatTok{1.2}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) +}\StringTok{ }\KeywordTok{xlab}\NormalTok{(}\StringTok{"stock size"}\NormalTok{) +}\StringTok{ }\KeywordTok{ylab}\NormalTok{(}\StringTok{"escapement"}\NormalTok{)  +}
\StringTok{  }\KeywordTok{scale_colour_manual}\NormalTok{(}\DataTypeTok{values=}\NormalTok{colorkey)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/nonparametric-bayes-Figure2.pdf}
\caption{The steady-state optimal policy (infinite boundary) calculated
under each model. Policies are shown in terms of target escapement,
$S_t$, as under models such as this a constant escapement policy is
expected to be optimal (Reed 1979).}
\end{figure}

The resulting optimal management strategy based on each of the inferred
models is shown in Figure 2, against the optimal strategy given the true
underlying dynamics. Policies are shown in terms of target escapement,
$S_t$. Under models such as this a constant escapement policy is
expected to be optimal (Reed 1979), whereby population levels below a
certain size $S$ are unharvested, while above that size the harvest
strategy aims to return the population to $S$, resulting in the
hockey-stick shaped policies shown.
